\sysfull

### 2017-08-09 paper/compiler.tex
Compiler section revised

\sys provides compiler support for \emph{automatic task generation} that simplifies writing task-based code by automatically translating C code into a task-based executable.  Translating C code into task-based code automatically requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime. The \sys compiler first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary. After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.

@@

The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}. The \sys compiler builds the program's CFG (as shown in
Figure~\ref{fig:comp_2}). To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG. \sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops. We first describe how \sys measures each of these costs and then describe \sys's procedure for placing boundaries in the CFG.

@@

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block. Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we used the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost is likely to closely correspond to block length. Moreover, we
assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy. This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement. \sys's analysis tries ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget.

{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of {\em live variables} that are accessible in both of the blocks that define the edge (i.e., predecessor and successor). Variables that live across an edge between two tasks must be protected by \sys's memory virtualization mechanism and each live, inter-task variable increases the run time overhead for a boundary placement. \sys's analysis tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.

{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
annotated with a static loop bound, which is common in embedded code. For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.

@@

The analysis uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score. The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of variables live variables at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across applications. We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.

>>>>

\sys provides compiler support for \emph{automatic task generation} that simplifies writing task-based code by automatically translating C code into a task-based executable. Doing so requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime. The \sys compiler, \textcolor{red}{based on LLVM}, first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary. After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.

@@

The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}. The \sys compiler builds the program's CFG (as shown in
Figure~\ref{fig:comp_2}). To simplify inter-procedural analysis, our prototype inlines all function calls\todo{how about recursions and code duplication?}{Brandon/Kiwan} into the main function in the CFG.
%
%\sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops.
%
We first describe how \sys measures each of these costs and then describe \sys's procedure for placing boundaries in the CFG.

@@

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block. Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost. Moreover, we assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy. This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement. \sys's analysis tries to ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget (which is part of the boundary placement process described later).

{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of {\em live variables} that are accessed in both\todo{Transitive closure?}{Brandon/Kiwan} of the blocks that define the edge (i.e., predecessor and successor). Variables that live across an edge between two tasks must be protected by \sys's memory virtualization mechanism and each live, inter-task variable increases the run time overhead for a boundary placement. \sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.

{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.

%removed statement "which is common in embedded code"

@@

\sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score. The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of variables live variables at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across applications. We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.
